---
title: 'Stat 480 - Homework #5'
author: "Jeremy Miranda"
date: "2/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

<br/>

#### Submission Details

**Due date**: the homework is due before class on Thursday. 

**Submission process**: submit both the R Markdown file and the corresponding html file on canvas. Please submit both the `.Rmd` and the `.html` files separately and do not zip the two files together.

<br/>

## Flying etiquette

1. Download the RMarkdown file with these homework instructions to use as a template for your work.
Make sure to replace "Your Name" in the YAML with your name.


2. FiveThirtyEight is a website founded by Statistician and writer Nate Silver to publish results from  opinion poll analysis, politics, economics, and sports blogging. One of the featured articles considers [flying etiquette](https://fivethirtyeight.com/features/airplane-etiquette-recline-seat/). This article is based on data collected by FiveThirtyEight and publicly available on github. Use the code below to read in the data from the survey:

```{r}
fly <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/flying-etiquette-survey/flying-etiquette.csv")
```

The next couple of lines of code provide a bit of cleanup of the demographic information by reordering the levels of the  corresponding factor variables. Run this code in your session.

```{r}
fly$Age <- factor(fly$Age, levels=c("18-29", "30-44", "45-60", "> 60", ""))
fly$Household.Income <- factor(fly$Household.Income, levels = c("$0 - $24,999","$25,000 - $49,999", "$50,000 - $99,999", "$100,000 - $149,999", "150000", ""))
fly$Education <- factor(fly$Education, levels = c("Less than high school degree", "High school degree", "Some college or Associate degree", "Bachelor degree",  "Graduate degree", ""))
```


3. Some people do not travel often by plane. Provide a breakdown of travel frequency (use variable `How.often.do.you.travel.by.plane.`). Reorder the levels in the variable by travel frequency from least frequent travel to most frequent. Draw a barchart of travel frequency and comment on it. 

```{r}
library(forcats)
library(ggplot2)
ggplot(fly, aes(x = fct_infreq(fly$How.often.do.you.travel.by.plane.))) + 
  geom_bar() + 
  coord_flip()
```


## Interpretation

As the bar chart shows most people fly one time per year or less, and it's really not even close. Once a month or less is the 2nd closest but even that option is around 400 counts away.

4. In the demographic variables (`Education`, `Age`, and `Houshold.Income`), replace all occurrences of the empty string "" by a missing value `NA`. How many responses do not have any missing values? (Hint: the function `is.na` might come in handy)

```{r}
library(dplyr)
levels(fly$Education)
levels(fly$Education)[6] <- NA
summary(is.na <- (fly$Education))

levels(fly$Age)
levels(fly$Age)[5] <- NA
summary(is.na <- (fly$Age))

levels(fly$Household.Income)
levels(fly$Household.Income)[6] <- NA
summary(is.na <- (fly$Household.Income))
```


## Interpretaion

Number of NA's for education is 39
Number of NA's for Age is 33
Number of NA's for household income is 214


5. Run the command below and interpret the output. What potential purpose can you see for the chart? 
```{r, fig.width=8, fig.height=3, message=FALSE}
library(ggplot2)
fly$Education = with(fly, factor(Education, levels = rev(levels(Education))))

ggplot(data = fly, aes(x = 1)) + 
  geom_bar(aes(fill=Education), position="fill") + 
  coord_flip() +
  theme(legend.position="bottom") +
  scale_fill_brewer() + 
  xlab("Ratio") 
```


## Interpretation

This chart can be used to better compare the different levels of education by providing a proportin/percentage breakdown of each level.


6. Rename the variable `In.general..is.itrude.to.bring.a.baby.on.a.plane.` to `baby.on.plane.`. How many levels does the variable `baby.on.plane` have, and what are these levels? Rename the level labeled "" to "Not answered". Reorder the levels of `baby.on.plane` from least rude to most rude. Put the level "Not answered" last. Draw a barchart of variable `baby.on.plane`.  Interpret the result.

```{r}
library(dplyr)
fly2 <- rename(fly, baby.on.plane = In.general..is.itrude.to.bring.a.baby.on.a.plane.)
fly <- fly2

levels(fly$baby.on.plane)
levels(fly$baby.on.plane)[1] <- "Not Answered"

levels(factor(fly$baby.on.plane, levels = c("No, not at all rude", "Yes somewhat rude", "Yes, very rude", "Not Answered")))

library(ggplot2)
ggplot(fly, aes(x = fly$baby.on.plane)) + 
  geom_bar() + 
  coord_flip()
```


## Interpretation

The bar chart shows that most people don't view bringing a baby on a plane as being rude. The result doesn't surprise me people understand that everyone has to travel and if you have a baby you can't leave your baby at home.


7. Investigate the relationship between gender and the variables `Do.you.have.any.children.under.18.` and `baby.on.plane`. How is the attitude towards babies on planes shaped by gender and having children under 18? Find a plot that summarises your findings (use `ggplot2`).
```{r, fig.width=30, fig.height=25}
library(ggmosaic)

ggplot(data = fly)  + 
  geom_mosaic(aes(x = product(Gender, baby.on.plane), fill = fly$Do.you.have.any.children.under.18., weight=1)) + theme(axis.text.x = element_text(size = 15, angle = 90), axis.text.y = element_text(size = 20)) + facet_grid(fly$baby.on.plane)
```


## Interpretation

Full disclosure I don't actual know if I did this question right. I was a bit confused because I wasn't sure if I needed two mosaic plots or one. I was able to figure out how to get all the variables into one plot so I figured that would be the right choice. The mosaic plot shows that whether the person asked has any children under the age of 18 or not or the person is male of female the answer to the baby on plane question based on the previously mentioned stipulations is the same.


